---
title: "MLE To Bayes Factors"
author: "Bradley T. Martin"
date: "April 4, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# BFD MLE to Bayes Factors  

This is a workflow to generate Bayes Factors from marginal likelihood estimates (MLE) obtained via BFD/SNAPP and BEAST2 Path Sampling. 


## Input BFD Data  

Input: The input is a CSV file containing the following columns (order does not matter):  

1. "MLE"
  + Can have NA values; they will be excluded.  
  
2. "rank"
  + Should be blank (NA).  
  
3. "BFD_run"
  + Can be integer or string.  
  
4. "model"
  + Can be integer or string.  
  
5. "BF"
  + Should be blank (NA).  


First, load appropriate libraries:  

```{r load.libraries, results="hide"}
library("knitr")
library("kableExtra")
```

## Loading and Summarizing the CSV file.  

```{r csv2df}
file <- file.choose()
df <- read.csv(file = file, header = TRUE, sep = ",", stringsAsFactors = FALSE)
```


Produce a summary table of the ESS (Effective Samples Sizes) for each model:  

```{r summarizeDF}

# Combine BFD_run and model columns for better 
# display. 
df["BFD Model"] <- 
  paste0("run", 
         df[, "BFD_run"], 
         ", ", 
         df[, "model"])

# Combine Median, Min, and Max columns for better
# display. 
df["MLE Med (Min, Max)"] <- 
  paste0(df[, "MLE_ESS_median"], 
         " (", 
         df[, "MLE_ESS_min"], 
         ", ", 
         df[, "MLE_ESS_max"], ")")

# Get the summary columns into a new DataFrame.
run.summary <- df[, c("BFD Model", 
                      "MLE_ESS_avg", 
                      "MLE Med (Min, Max)", 
                      "MLE", "MLE_SD")]

# Rename the summary columns for Table
colnames(run.summary) <- 
  c("BFD Model", 
    "ESS Mean", 
    "ESS Median (Min, Max)", 
    "MLE", "Std. Dev.")

# Make the table using knitr::kable.
kable(run.summary, 
      format = "latex", 
      align = "l", 
      row.names = FALSE, 
      digits = 2, 
      caption = "BFD* ESS Summary", 
      booktabs = TRUE) %>% 
  
  kable_styling(latex_options = 
                  c("scale_down",
                  "hold_position")) %>%
  
  row_spec(row = 0,
           bold=TRUE)
```

Now remove missing data for runs that haven't finished yet:  

## Remove missing data


```{r dropNA}
# This removes rows with NA in the MLE column.  
completeVec <- 
  complete.cases(df[, "MLE"])

# Make a new dataframe without unfinished BFD runs.  
df2 <- df[completeVec, ]
```



## Calculating the Bayes Factors

The below chunk of code first finds the best model based on having the higest MLE. Then it ranks them and caluculates Bayes Factors for each one like so:    
BF = (2 * (model1 - model2))  

A negative value indicates support for model 1.  
A positive value indicates support for model 2.  

Support for the models is as shown from Kass and Raftery (1995) and Adam Leache's BFD Tutorial:  

0 < BF < 2 is not worth more than a bare mention  
2 < BF < 6 is positive evidence   
6 < BF < 10 is strong support   
and BF > 10 is decisive   

I made model1 the allSplit model (run20) for all comparisons.  


```{r calcBF}

# Rank the MLE's. Highest = best model.
df2$rank <- rank(x = abs(df2$MLE))

# Get the MLE of rank 1.
for (i in 1:nrow(df2)) { 
  if (df2$rank[i] == 1) 
  { 
    best.mle <- df2$MLE[i] 
  } 
}

# calculate the Bayes Factors.
df2[, "BF"] <- (2 * (best.mle - df2[, "MLE"]))

# Set BF of rank 1 to NA.
for (i in 1:nrow(df2)) {
  if (df2$rank[i] == 1)
  {
    df2$BF[i] <- NA
  }
}

# Sort the dataframe rows by rank column.
df2 <- df2[order(df2$rank), ]

```



Now make a nice table for the BF data:  



```{r makeBFDtable}

# Make the NA go away when comparing model1 to
# itself.
options(knitr.kable.NA = "-")

# Just get columns of interest.
bfd.tabdf <- 
  df2[, c("BFD Model", 
          "MLE", 
          "K",
          "rank",
          "BF")]

# Rename columns.
colnames(bfd.tabdf) <- 
  c("BFD Model", 
    "MLE", 
    "K",
    "Rank",
    "BF")

# Write to table.
kable(bfd.tabdf, 
      format = "latex", 
      booktabs = TRUE, 
      digits = 2, 
      align = c("l", "c", "c", "c"), 
      row.names = FALSE, 
      caption = "BFD* Bayes Factors") %>%
  
  kable_styling(full_width = TRUE, 
                latex_options = 
                  "hold_position") %>%
  
  column_spec(column = 1, 
              width = "9cm") %>%
 
   row_spec(row = 0,
           bold=TRUE)
```

Now it's done!